---
title: "ST558Project1"
author: "Yi Ren"
date: '2022-09-06'
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(ggplot2)
library(tidyverse)
library(dplyr)
```

# Create multiple functions
```{r function, include = TRUE}
filtdata <- function(url, var_name = "value"){
  sheet <- read_csv(url) 
  sh <- sheet %>% select(Area_name, STCOU, ends_with("D")) %>% rename(area_name=`Area_name`) 
  return(sh)
}
convert <- function(sh){
  sh_longer <- pivot_longer(sh, cols = 3:12, names_to = "name", values_to = "value")
  sh_longer$measurements <- substr(sh_longer$name, 1, 7) 
  sh_longer$year <- as.numeric(substr(sh_longer$name, 8, 9))
  sh_longer$year <- ifelse(sh_longer$year > 50, sh_longer$year + 1900, sh_longer$year + 2000) 
  return(sh_longer)
}
disstat <- function(sh_longer){
  county <- sh_longer[grep(pattern = ", \\w\\w", sh_longer$area_name), ]
  class(county) <- c("county", class(county))
  n_last <- 2
  county$state <-substr(county$area_name, nchar(county$area_name) - n_last + 1, nchar(county$area_name))
  return(county)
}
disdivi <- function(sh_longer){
  non_county <- sh_longer[grep(pattern = ", \\w\\w", sh_longer$area_name, invert = TRUE), ]
  class(non_county) <- c("state", class(non_county))
  non_county[ , "division"] <- NA
  non_county$division[non_county$area_name %in% c("UNITED STATES")]="ERROR"
  non_county$division[non_county$area_name %in% c("CONNECTICUT", "MAINE", "MASSACHUSETTS", "NEW HAMPSHIRE", "RHODE ISLAND", "VERMONT")]="New England"
  non_county$division[non_county$area_name %in% c("NEW JERSEY", "NEW YORK", "PENNSYLVANIA")]="Mid-Atlantic"
  non_county$division[non_county$area_name %in% c("ILLINOIS", "INDIANA", "MICHIGAN", "OHIO", "WISCONSIN")]="East North Central"
  non_county$division[non_county$area_name %in% c("IOWA", "KANSAS", "MINNESOTA", "MISSOURI", "NEBRASKA", "NORTH DAKOTA", "SOUTH DAKOTA")]="West North Central"
  non_county$division[non_county$area_name %in% c("DELAWARE", "FLORIDA", "GEORGIA", "MARYLAND", "NORTH CAROLINA", "SOUTH CAROLINA", "VIRGINIA","DISTRICT OF COLUMBIA", "District of Columbia", "WEST VIRGINIA")]="South Atlantic"
  non_county$division[non_county$area_name %in% c("ALABAMA", "KENTUCKY", "MISSISSIPPI", "TENNESSEE")]="East South Central"
  non_county$division[non_county$area_name %in% c("ARKANSAS", "LOUISIANA", "OKLAHOMA", "TEXAS")]="West South Central"
  non_county$division[non_county$area_name %in% c("ARIZONA", "COLORADO", "IDAHO", "MONTANA", "NEVADA", "NEW MEXICO", "UTAH", "WYOMING")]="Mountain"
  non_county$division[non_county$area_name %in% c("ALASKA", "CALIFORNIA", "HAWAII", "OREGON", "WASHINGTON")]="Pacific"
  return(non_county)
}
```

# Create wrapper
```{r wrapper, include = TRUE}
my_wrapper <- function(url, var_name = "value"){
  a <- filtdata (url)
  b <- convert (a)
  c <- disstat (b)
  d <- disdivi (b)
  return(list(c, d))
}
```

# Read in data and combine them approperately
```{r df, include = TRUE}
dfa <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/EDU01a.csv")
dfb <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/EDU01b.csv")
county_df <- dplyr::bind_rows(dfa[[1]], dfb[[1]])
state_df <- dplyr::bind_rows(dfa[[2]], dfb[[2]])
```

# Create function for state plot
```{r stplot, echo = TRUE}
plot.state <- function(df, var_name = "value"){
  avg <- df %>% group_by(division, year) %>% summarise(mean = mean(get(var_name))) %>% filter(division != "ERROR")
  ggplot(avg, aes(x = year, y = mean, color = division)) + geom_line()
}
```

# Create function for county plot
```{r plot, echo = TRUE}
plot.county <- function(f, var_name = "value", state = "AL", top = TRUE, num_top = 5){
  df <- f %>% filter(state == !!state) %>% select(area_name, value)
  st_df <- summarise_all(group_by(df, area_name), mean)
  col <- st_df %>% pull(get(var_name))
  if(!top) {
    or_df <- st_df[order(col), ]
  } else {
    or_df <- st_df[order(col, decreasing = TRUE), ]
}
  fi_df <- head(or_df, num_top)
  value_df <- f[f$area_name %in% c(fi_df$area_name), ]
  ggplot(value_df, aes(x = year, y = value, color = area_name)) + geom_line()
}
```

## Generate required plot for state
```{r stateplot, echo = TRUE}
plot(state_df, var_name = "value")
```

## Generate required plot for county
```{r countyplot, echo = TRUE}
#Plot 1
plot(county_df, var_name = "value", state = "PA", top = TRUE, num_top = 7) 
#Plot 2
plot(county_df, var_name = "value", state = "PA", top = FALSE, num_top = 4)
#Plot 3
plot(county_df)
#Plot 4
plot(county_df, var_name = "value", state = "MN", top = TRUE, num_top = 10) 
```

# Read in new daataset using same functions
```{r newsheet, include = TRUE}
newa <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01a.csv")
newb <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01b.csv")
newc <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01c.csv")
newd<- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01d.csv")

county_ab <- dplyr::bind_rows(newa[[1]], newb[[1]])
county_cd <- dplyr::bind_rows(newc[[1]], newd[[1]])
county_new <- dplyr::bind_rows(county_ab, county_cd)

state_ab <- dplyr::bind_rows(newa[[2]], newb[[2]])
state_cd <- dplyr::bind_rows(newc[[2]], newd[[2]])
state_new <- dplyr::bind_rows(state_ab, state_cd)
```

## Plot new dataset for state
```{r newstateplot, echo = TRUE}
plot(state_new, var_name = "value")
```

## Plot new dataset for county
```{r newcoutyplot, echo = TRUE}
#Plot 1
plot(county_new, var_name = "value", state = "CT", top = TRUE, num_top = 6) 
#Plot 2
plot(county_new, var_name = "value", state = "NC", top = FALSE, num_top = 10)
#Plot 3
plot(county_new)
#Plot 4
plot(county_new, var_name = "value", state = "MN", top = TRUE, num_top = 4) 